Deep Recurrent Neural Networks for Supernovae Classification

### 2016-10-31 paper/sn.tex
LICENSE added
Rewording paper

Using post SPCC data, we obtain results which are competitive with previous approaches.  Although we have trained the network on the cross-entropy loss  and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65). Recurrent neural networks  therefore compare very well with a larger training  set. The performance isn't  quite as good with a smaller training set, possibly due to the network having to learn from no prior information about (noisy) light curves. This is also seen in~\cite{Lochner:2016hbn}, which obtains improved metrics for only a small training fraction with a combination of feature extraction and machine learning. As well as finding competitive results for the final metrics, we have shown that it is possible to give fast, early evaluation of supernovae type using pre-trained models. This is possible since the light curve can be fed to the model directly without needing any feature extraction.

>>>>

Using post SPCC data, we obtain results which are competitive with previous approaches.  Although we have trained the network on the cross-entropy loss  and not the $F_1$ score, for the same sized dataset of $\sim10^3 (10^4)$ supernovae (including host galaxy photo-z), \cite{Karpenka:2012pm} obtained $F_1$ values of 0.33 (0.58), and \cite{Newling:2010bp} values of 0.42 (0.57), compared to our  0.31 (0.65). Recurrent neural networks  therefore compare well with other methods when a larger training  set is available. The performance isn't  quite as good with a smaller training set, possibly due to the network having to learn from no prior information about (noisy) light curves. The current state of the art for a small training set  ($\sim10^3$ supernovae) comes from a combination of SALT2 (Spectral Adaptive Light curve Template 2) template fits and boosted decision trees~\cite{Lochner:2016hbn}. It would be interesting to check how how deep learning compares to this with a larger training set.

As well as finding competitive results for the final metrics, we have shown that it is possible to give fast, early evaluation of supernovae type using pre-trained models. This is possible since the light curve can be fed to the model directly without needing any feature extraction.
